Based on polling as of 27 October 2024, Harris expects to win on 50%, leads Trump by about 3% in 2024 US election

### 2024-10-30 paper/paper.qmd
Update paper.qmd

---
title: "Based on polling as of 27 October 2024, Harris expects to win on xxx%, leads Trump by xx% in 2024 US election"
subtitle: "Weighted Linear Regression Analysis of Aggregated Poll Data: Utilizing National Polls indicator, Population, and Pollster as Predictors."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

The data used to predict the outcome of the 2024 U.S. presidential
election in this study came from a variety of polling sources. The
results of these polls represent key measures of electoral support and
voter sentiment, and are therefore the main variables in our analysis.

To ensure that polling data accurately reflect real-world voter
preferences, we rely on several key measurement constructs. First, voter
preferences are measured by the percentage of respondents who indicated
support for each candidate. Each poll is conducted using different
sample sizes, sampling methods (e.g., online surveys, IVR calls, etc.),
which can inject more variability into the data. To mitigate these
differences, we aggregate poll results using a “”poll of polls“”
methodology that eliminates anomalies and provides a more comprehensive
picture of voter preferences (Blumenthal 2014; Pasek 2015).

@@

We conducted a regression analysis to predict support for Kamala Harris
and Donald Trump in the 2024 U.S. presidential election, including
models for Trump for comparison. Our analysis includes four models:
**unweighted linear models** and **weighted linear models** for both
candidates. These models predict each candidate’s support as a
continuous outcome across different states, using **pollster**,
**national poll**, and **population** as key predictors. The weighted
models additionally incorporate adjustments for recency, sample size,
and poll quality to account for potential biases. This setup aligns with
the New York Times methodology for election polling, which assigns
weights based on poll quality, sample size, and recency to better
reflect voter sentiment [@nytimes].

@@

While this analysis employs linear regression with weighted least
squares, alternative methods such as Bayesian modeling could provide
additional insights. However, we opted for the weighted linear
regression approach to maintain interpretability and ease of
implementation given the scope and resources of this analysis. The
Bayesian approach was not chosen primarily due to the complexity
involved in defining priors and the additional computational
requirements.

@@

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

@@

>>>>

---
title: "Based on polling as of 27 October 2024, Harris expects to win on 50%, leads Trump by about 3% in 2024 US election"
subtitle: "Weighted Linear Regression Analysis of Aggregated Poll Data: Utilizing National Polls indicator, Population, and Pollster as Predictors."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

The data used to predict the outcome of the 2024 U.S. presidential
election in this study came from a variety of polling sources. The
results of these polls represent key measures of electoral support and
voter sentiment and are therefore the main variables in our analysis.

To ensure that polling data accurately reflects real-world voter
preferences, we rely on several key measurement constructs. First, voter
preferences are measured by the percentage of respondents who indicated
support for each candidate. Each poll is conducted using different
sample sizes, and sampling methods (e.g., online surveys, IVR calls, etc.),
which can inject more variability into the data. To mitigate these
differences, we aggregate poll results using a “”poll of polls“”
methodology that eliminates anomalies and provides a more comprehensive
picture of voter preferences (Blumenthal 2014; Pasek 2015).

@@

We conducted a regression analysis to predict support for Kamala Harris
and Donald Trump in the 2024 U.S. presidential election, including
models for Trump for comparison. Our analysis includes four models:
**unweighted linear models** and **weighted linear models** for both
candidates. These models predict each candidate’s support as a
continuous outcome across different states, using **pollster**,
**national poll**, and **population** as key predictors. The weighted
models additionally incorporate adjustments for recency, sample size,
and poll quality to account for potential biases. This setup aligns with
the New York Times methodology for election polling, which assigns
weights based on poll quality, sample size, and recency to better
reflect voter sentiment [@nyt_polling_averages].

@@

While this analysis employs linear regression with weighted least
squares, alternative methods such as Bayesian modeling could provide
additional insights. However, we opted for the weighted linear
a regression approach to maintain interpretability and ease of
implementation given the scope and resources of this analysis. The
Bayesian approach was not chosen primarily due to the complexity
involved in defining priors and the additional computational
requirements.

@@

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reached just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

@@

## Recent 3 Months Support Trends Linear Model Prediction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
set.seed(233)
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  national_poll = rep(c(0, 1), length(1000)),
  pollster = factor(sample(levels(harris_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(harris_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
  sample_size = sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = pollster_quality / 4,

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed
        recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1),
    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Harris"
  )
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 1000
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(trump_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
   national_poll = rep(c(0, 1), length(1000)),  
  sample_size =sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = pollster_quality / 4,

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1),

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-predicted
#| fig-cap: Predicted Support Trends for Kamala Harris and Donald Trump using Fitted Linear Model
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

# Predictions for Trump
new_data_trump <- new_data_trump |>
  mutate(
    predicted_unweighted = predict(trump_unweighted_model, newdata = new_data_trump),
    predicted_weighted = predict(trump_weighted_model, newdata = new_data_trump)
  )

# Define the cutoff date for recent data
recent_date_cutoff <- as.Date("2024-08-01")

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- max(combined_predictions_filtered$end_date)

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = TRUE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = TRUE, size = 1) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Weights: recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight\nLinear Regression Model: candidate support percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

The predicted support for Kamala Harris and Donald Trump in the upcoming 2024 U.S. presidential election is illustrated in Figure 1. This analysis spans the period from August 1, 2024, to October 27, 2024, and uses multiple linear regression models to project voter support.

The unweighted model predictions indicate that Kamala Harris maintains a consistent support level, with projections hovering around 49%. The blue dashed line represents this trend, showing only minor fluctuations throughout the polling period. Conversely, the weighted model for Harris, indicated by the solid blue line, suggests a slight upward trend in predicted support, reflecting a potential positive shift in voter sentiment as the election date approaches.

In contrast, Donald Trump's predicted support is represented by the red lines. The unweighted model forecasts his support to remain around 48%, while the weighted model predicts a gradual decline to approximately 46%. The red dashed line indicates the unweighted prediction, whereas the solid red line shows the weighted prediction. This downward trajectory may indicate challenges for Trump in sustaining voter enthusiasm amid changing public opinions.

At the conclusion of the polling period, the models suggest that Kamala Harris leads Donald Trump by approximately 4% in the un-weighted (50% vs. 46%) and 2.5% in weighted models (50% vs. 47.5%).
